raptr: Representative and Adequate Prioritization Toolkit in R

### 2016-11-23 code/rmarkdown/supporting-information.Rmd
+ add in readmes + compile manuscript into single pdf and single docx

---
fontsize: 11pt
documentclass: article
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble.tex, si-preamble.tex]
---
# Supporting Information

@@

@@

## Appendix S1
The reliable formulation explicitly considers the probability that planning units are inhabited by features. As a consequence, it may deliver prioritisations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritisation is implemented. This behaviour is achieved by siting back-up planning units near selected planning units with low occupancy probabilities in the attribute space(s). To ensure that prioritisations are robust against multiple planning units being uninhabited, the problem assigns planning units at multiple backup levels.

Backup levels levels are defined as $r$-levels [similar to failure levels in @r18]. The first backup $r$-level is used to calculate the level of representation when all of the selected planning units are occupied by all $f \in F$. For this scenario, the closest selected planning unit to each demand point $i$ for attribute space $s$ is assigned at $r$-level$=0$. This scenario essentially represents $Y_{fsij}$ in the unreliable formulation. The second backup $r$-level is used to assess the level of representation when the closest planning unit to each demand point $i$ is unoccupied. For this scenario, the second closest planning units are assigned at $r$-level$=1$. The third backup $r$-level is used to assess representation when the first two closest planning units are unoccupied. The third closest planning units are assigned at $r$-level$=2$. Continuing on, in this manner, the selected planning units in a prioritisation are assigned to each demand point $i \in I$, attribute space $s \in S$, and each feature $f \in F$ at an $r$-level.

@@

Moderately-sized conservation planning problems often include several thousand planning units. It is currently not be feasible to solve this problem when considering all possible failure scenarios. As a consequence, the $R$ variable can be any $1 \leq R \leq J-1$. For instance, when $R=3$ only 2 backup levels are considered in addition to the final backup level. Cui et al. [-@r16] found that $R=5$ yields similar solutions to $R >> 5$. However, in most cases the decision maker will likely be limited to $R=1$ to obtain prioritisations in a feasible amount of time.

@@

The reliable formulation is non-linear. However, the non-linear components can be linearised. First--as discussed in the main text--the expression $X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer et al. [-@r426]. Second, the expression $P_{fsijr} Y_{fsijr}$ in (eqn 7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al. [-@r16].

## Figures 
```{r, echo=FALSE, fig.height=8, fig.width=8.5, fig.cap='Attribute spaces used in the first case-study. Each panel shows a the distribution of a solution in environmental space and how it samples the realised niche for a different species. The left column of panels shows the solution generated using amount targets. The right column of panels shows the solution generated using amount and niche targets. Each column of panels corresponds to a different species. Hexagons show the distribution of demand points. The color of each heaxgon denotes the weighted frequency of demand points inside it. Points represent the environmental conditions associated with planning units inside the species geographic range that were selected for preservation in a given solution.'}
## prepare data
cs1.pu.niche.DF <- ddply(
	cs1.spp.DF,
	c('Species', 'Prioritisation'),
	function(x) {
		# init
		curr.prioritisation <- cs1.prioritisations[[as.character(x$Prioritisation)]]
		curr.spp.pos <- match(x$Species[1], curr.prioritisation@data@species[[1]])
		curr.sel.pos <- which(curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@ids %in% which(as.logical(selections(curr.prioritisation))))
		# extract planning units
		curr.DF <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@coords %>% as.data.frame()
		names(curr.DF) <- c('Niche1', 'Niche2')
		curr.DF$Species <-x$Species[[1]]
		curr.DF$Prioritisation <-x$Prioritisation[[1]]
		curr.DF$Solution <- 'discarded'
		curr.DF$Solution[curr.sel.pos] <- 'selected'
		# return data
		return(curr.DF)
	}
)
cs1.pu.niche.DF <- filter(cs1.pu.niche.DF, Solution=='selected')

@@

```{r, echo=FALSE, results='asis'}
# make table
term.CHR <- c(
	'$F$', 'set of biodiversity features (indexed by $f$)',
	'$J$', 'set of planning units (indexed by $j$)',
	'$q_{fj}$', 'probability of feature $f$ occupying planning unit $j$',
	'$e_{jk}$', 'shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrise exposed edges with no neighbours.',
	'$S$', 'set of attribute spaces (indexed by $S$)',
	'$I_{fsi}$', 'set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$',
	'$\\lambda_{fsi}$', 'set of weights for demand point $i$ for feature $f$ in attribute space $s$',
	'$d_{fsij}$', 'distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$',
	'$\\delta_{fsi}$', 'the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$',
	'$T_f$', 'amount target for feature $f$',
	'$\\tau_{fs}$', 'space-based target for feature $f$ in attribute space $s$',
	'$X_j$', 'binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0)',
	'$Y_{fsij}$', 'binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units'
)
term.DF <- data.frame(
	Symbol=term.CHR[seq(1, length(term.CHR), 2)],
	Description=term.CHR[seq(2, length(term.CHR), 2)]
)
term.DF$Description <- paste0('\\parbox[t][][t]{15cm}{', term.DF$Description, '}')

# render table
curr.file <- tempfile(fileext='.tex')
tmp <- latex(
	term.DF,
	file=curr.file, 
	digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
	colheads=c('Symbol', 'Description'), col.just=c('c', 'l'),
	caption='Symbols and descriptions of terms used in the formulation of the unreliable representative and adequate prioritisation (URAP) problem.'
)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```

# References

>>>>

---
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
---
# Supporting Information

@@

\setcounter{figure}{0}
\setcounter{table}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

@@

## Appendix S1
The reliable formulation explicitly considers the probability that planning units are inhabited by features. As a consequence, it may deliver prioritizations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritization is implemented. This behavior is achieved by siting back-up planning units near selected planning units with low occupancy probabilities in the attribute space(s). To ensure that prioritizations are robust against multiple planning units being uninhabited, the problem assigns planning units at multiple backup levels.

Backup levels levels are defined as $r$-levels [similar to failure levels in @r18]. The first backup $r$-level is used to calculate the level of representation when all of the selected planning units are occupied by all $f \in F$. For this scenario, the closest selected planning unit to each demand point $i$ for attribute space $s$ is assigned at $r$-level$=0$. This scenario essentially represents $Y_{fsij}$ in the unreliable formulation. The second backup $r$-level is used to assess the level of representation when the closest planning unit to each demand point $i$ is unoccupied. For this scenario, the second closest planning units are assigned at $r$-level$=1$. The third backup $r$-level is used to assess representation when the first two closest planning units are unoccupied. The third closest planning units are assigned at $r$-level$=2$. Continuing on, in this manner, the selected planning units in a prioritization are assigned to each demand point $i \in I$, attribute space $s \in S$, and each feature $f \in F$ at an $r$-level.

@@

Moderately-sized conservation planning problems often include several thousand planning units. It is currently not be feasible to solve this problem when considering all possible failure scenarios. As a consequence, the $R$ variable can be any $1 \leq R \leq J-1$. For instance, when $R=3$ only 2 backup levels are considered in addition to the final backup level. Cui et al. [-@r16] found that $R=5$ yields similar solutions to $R >> 5$. However, in most cases the decision maker will likely be limited to $R=1$ to obtain prioritizations in a feasible amount of time.

@@

The reliable formulation is non-linear. However, the non-linear components can be linearized. First--as discussed in the main text--the expression $X_j X_k$ in (eqn 3a) can be linearized using methods described by Beyer et al. [-@r426]. Second, the expression $P_{fsijr} Y_{fsijr}$ in (eqn 7a) can be linearized using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al. [-@r16].

## Figures 
```{r, echo=FALSE, fig.height=8, fig.width=8.5, fig.cap='Attribute spaces used in the first case-study. Each panel shows a the distribution of a solution in environmental space and how it samples the realized niche for a different species. The left column of panels shows the solution generated using amount targets. The right column of panels shows the solution generated using amount and niche targets. Each column of panels corresponds to a different species. Hexagons show the distribution of demand points. The color of each heaxgon denotes the weighted frequency of demand points inside it. Points represent the environmental conditions associated with planning units inside the species geographic range that were selected for preservation in a given solution.'}
## prepare data
cs1.pu.niche.DF <- ddply(
	cs1.spp.DF,
	c('Species', 'Prioritisation'),
	function(x) {
		# init
		curr.prioritisation <- cs1.prioritisations[[as.character(x$Prioritisation)]]
		curr.spp.pos <- match(x$Species[1], curr.prioritisation@data@species[[1]])
		curr.sel.pos <- which(curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@ids %in% which(as.logical(selections(curr.prioritisation))))
		# extract planning units
		curr.DF <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@coords %>% as.data.frame()
		names(curr.DF) <- c('Niche1', 'Niche2')
		curr.DF$Species <-x$Species[[1]]
		curr.DF$Prioritisation <-x$Prioritisation[[1]]
		curr.DF$Solution <- 'discarded'
		curr.DF$Solution[curr.sel.pos] <- 'selected'
		# return data
		return(curr.DF)
	}
)
cs1.pu.niche.DF <- filter(cs1.pu.niche.DF, Solution=='selected')

@@

```{r, echo=FALSE, results='asis'}
# make table
term.CHR <- c(
	'$F$', 'set of biodiversity features (indexed by $f$)',
	'$J$', 'set of planning units (indexed by $j$)',
	'$q_{fj}$', 'probability of feature $f$ occupying planning unit $j$',
	'$e_{jk}$', 'shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrize exposed edges with no neighbours.',
	'$S$', 'set of attribute spaces (indexed by $S$)',
	'$I_{fsi}$', 'set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$',
	'$\\lambda_{fsi}$', 'set of weights for demand point $i$ for feature $f$ in attribute space $s$',
	'$d_{fsij}$', 'distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$',
	'$\\delta_{fsi}$', 'the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$',
	'$T_f$', 'amount target for feature $f$',
	'$\\tau_{fs}$', 'space-based target for feature $f$ in attribute space $s$',
	'$X_j$', 'binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0)',
	'$Y_{fsij}$', 'binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units'
)
term.DF <- data.frame(
	Symbol=term.CHR[seq(1, length(term.CHR), 2)],
	Description=term.CHR[seq(2, length(term.CHR), 2)]
)
term.DF$Description <- paste0('\\parbox[t][][t]{15cm}{', term.DF$Description, '}')

# render table
curr.file <- tempfile(fileext='.tex')
tmp <- latex(
	term.DF,
	file=curr.file, 
	digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
	colheads=c('Symbol', 'Description'), col.just=c('c', 'l'),
	caption='Symbols and descriptions of terms used in the formulation of the unreliable representative and adequate prioritization (URAP) problem.'
)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```

## References
